The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-03 rockerverse.Rmd
use \code{..} where appropriate, use DOI URLs
preparing for submission, #26

**\pkg{harbor}** ([https://github.com/wch/harbor](https://github.com/wch/harbor)) is no longer actively maintained, but it should be honorably mentioned as the first R package for managing Docker images and containers.
It uses the \CRANpkg{sys} package \citep{cran_sys} to run system commands against the Docker CLI, both locally and through an SSH connection, and it has convenience functions, e.g., for listing and removing containers/images and for accessing logs.
The outputs of container executions are converted to appropriate R types.
The Docker CLI's basic functionality, although it evolves quickly and with little concern for avoiding breaking changes, has remained unchanged in core functions, meaning that a core function such as `harbor::docker_run(image = "hello-world")` still works depsite its stopped development.

@@

**\CRANpkg{stevedore}** is currently the most powerful Docker client in R \citep{cran_stevedore}.
It interfaces with the Docker daemon over the Docker HTTP API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}} via a Unix socket on Linux or MacOS, over a named pipe on Windows, or over an HTTP/TCP connection.
The package is the only one not using system calls to the `docker` CLI tool for managing images and containers.
The package thereby enables connections to remote Docker instances without direct configuration of the local Docker daemon.
Furthermore using the API gives access to information in a structured way, is system independent, and is likely more reliable than parsing command line output.
\CRANpkg{stevedore}'s own interface is automatically generated based on the OpenAPI specification of the Docker daemon, but it is still similar to the Docker CLI.
The interface is similar to R6 objects, in that a `stevedore_object` has a number of functions attached to it that can be called, and multiple specific versions of the Docker API can be supported thanks to the automatic generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

@@

**\pkg{dockyard}** ([https://github.com/thebioengineer/dockyard](https://github.com/thebioengineer/dockyard)) has the goal of lowering the barrier to creating `Dockerfile`s, building Docker images, and deploying Docker containers.
The package follows the increasingly used piping paradigm of the `tidyverse` style of programming for chaining R functions representing the instructions in a `Dockerfile`.
An existing `Dockerfile` can be used as a template.
\pkg{dockyard} also includes wrappers for common steps, such as installing an R package or copying files, and provides built-in functions for building an image and running a container, which make Docker more approachable within a single R-based user interface.

@@

The _Bioconductor_ `Dockerfile` selects the desired version of R from Rocker, adds required system dependencies, and uses the \CRANpkg{BiocManager} package for installing appropriate versions of _Bioconductor_ packages \citep{cran_biocmanager}.
A strength of this approach is that the responsibility for complex software configuration (including customized development) is shifted from the user to the experienced _Bioconductor_ core team.
However, a recent audit of the _Bioconductor_  image stack `Dockerfile` led to the deprecation of several community-maintained images, because the numerous specific images became too hard to understand, complex to maintain, and cumbersome to customise.
As part of the simplification, a recent innovation is to produce a `bioconductor_docker:devel` image to emulate the _Bioconductor_ nightly Linux build machine as closely as possible.
This image contains the build system environment variables and the _system dependencies_ needed to install and check almost all (1813 out of 1823) _Bioconductor_ software packages, and it saves users and package developers from managing these themselves.
Furthermore, the image is configured so that `.libPaths()` has `/usr/local/lib/R/host-site-library` as the first location.
Users mounting a location on the host file system to this location can persistently manage installed packages across Docker containers or image updates.
Many R users pursue flexible work flows tailored to particular analysis needs rather than standardized work flows.
The new `bioconductor_docker` image is well suited for this preference, while `bioconductor_docker:devel` provides developers with a test environment close to _Bioconductor_'s build system.

@@

**\pkg{containerit}** ([https://github.com/o2r-project/containerit/](https://github.com/o2r-project/containerit/)) attempts to take this one step further and includes these tools to automatically create a `Dockerfile` that can execute a given workflow \citep{nust_containerit_2019}.
\pkg{containerit} accepts R `sessionInfo` objects as input and provides helper functions to derive these from workflows, e.g., an R script or R Markdown document, by analysing the session state at the end of the workflow.
It relies on the \pkg{sysreqs} ([https://github.com/r-hub/sysreqs/](https://github.com/r-hub/sysreqs/)) package and it's mapping of package system dependencies to platform specific installation package names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \CRANpkg{stevedore} to streamline the user interaction and improve the created `Dockerfile`s, e.g., by running a container for the desired base image to extract the already available R packages.

@@

**\CRANpkg{liftr}** aims to solve the problem of persistent reproducible reporting in statistical computing based on the R Markdown format \citep{xie2018} for dynamic documents \citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for collaborative documents and large-scale platforms for processing documents.
\CRANpkg{liftr} makes the document the main and sole workflow control file and the only file that needs to be shared between collaborators for consistent environments, e.g. demonstrated in the DockFlow project (https://dockflow.org).
It introduces new fields to the R Markdown document header, allowing users to manually declare the dependencies, including versions, for rendering the document.
The package then generates a `Dockerfile` from this metadata and provides a utitility function to render the document inside a Docker container, i.e., `render_docker("foo.Rmd")`.
An RStudio addin even allows compilation of documents with the single push of a button.

@@

**\CRANpkg{googleComputeEngineR}** has the function `gce_vm_cluster()` to create clusters of 2 or more virtual machines, running multi-CPU architectures \citep{googleComputeEngineR_2019}.

@@

Furthermore, \CRANpkg{drake} workflows have been shown to use \CRANpkg{future} package's function `makeClusterPSOCK()` for sending parts of the workflow to a Docker image for execution\footnote{\href{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}} \citep[see package's function documentation;~][]{future_2020}.

@@

Another example is the package \CRANpkg{golem}, which makes an heavy use of `dockerfiler` when it comes to creating the `Dockerfile` for building and deploying production-grade Shiny applications \citep{cran_golem}.
\CRANpkg{googleComputeEngineR} enables quick deployments of key R services, such as RStudio and Shiny, onto cloud virtual machines (VMs) with Google Cloud Compute Engine \citep{googleComputeEngineR_2019}.
The package utilises `Dockerfile`s to move the labour of setting up those services from the user to a premade Docker image, which is configured and run in the cloud VM.
For example, by specifying the template `template="rstudio"` in functions `gce_vm_template()/gce_vm()` an up-to-date RStudio Server image is launched for development work, whereas specifying `template="rstudio-gpu"` will launch an RStudio Server image with a GPU attached, etc.

>>>>

**\pkg{harbor}** ([https://github.com/wch/harbor](https://github.com/wch/harbor)) is no longer actively maintained, but it should be honorably mentioned as the first R package for managing Docker images and containers.
It uses the \CRANpkg{sys} package \citep{cran_sys} to run system commands against the Docker CLI, both locally and through an SSH connection, and it has convenience functions, e.g., for listing and removing containers/images and for accessing logs.
The outputs of container executions are converted to appropriate R types.
The Docker CLI's basic functionality, although it evolves quickly and with little concern for avoiding breaking changes, has remained unchanged in core functions, meaning that a core function such as \code{harbor::docker\_run(image = "hello-world")} still works depsite its stopped development.

@@

**\CRANpkg{stevedore}** is currently the most powerful Docker client in R \citep{cran_stevedore}.
It interfaces with the Docker daemon over the Docker HTTP API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}} via a Unix socket on Linux or MacOS, over a named pipe on Windows, or over an HTTP/TCP connection.
The package is the only one not using system calls to the `docker` CLI tool for managing images and containers.
The package thereby enables connections to remote Docker instances without direct configuration of the local Docker daemon.
Furthermore using the API gives access to information in a structured way, is system independent, and is likely more reliable than parsing command line output.
\CRANpkg{stevedore}'s own interface is automatically generated based on the OpenAPI specification of the Docker daemon, but it is still similar to the Docker CLI.
The interface is similar to R6 objects, in that an object of class \code{"stevedore\_object"} has a number of functions attached to it that can be called, and multiple specific versions of the Docker API can be supported thanks to the automatic generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

@@

**\pkg{dockyard}** ([https://github.com/thebioengineer/dockyard](https://github.com/thebioengineer/dockyard)) has the goal of lowering the barrier to creating `Dockerfile`s, building Docker images, and deploying Docker containers.
The package follows the increasingly used piping paradigm of the Tidyverse-style \citep{wickham_welcome_2019} of programming for chaining R functions representing the instructions in a `Dockerfile`.
An existing `Dockerfile` can be used as a template.
\pkg{dockyard} also includes wrappers for common steps, such as installing an R package or copying files, and provides built-in functions for building an image and running a container, which make Docker more approachable within a single R-based user interface.

@@

The _Bioconductor_ `Dockerfile` selects the desired version of R from Rocker, adds required system dependencies, and uses the \CRANpkg{BiocManager} package for installing appropriate versions of _Bioconductor_ packages \citep{cran_biocmanager}.
A strength of this approach is that the responsibility for complex software configuration (including customized development) is shifted from the user to the experienced _Bioconductor_ core team.
However, a recent audit of the _Bioconductor_  image stack `Dockerfile` led to the deprecation of several community-maintained images, because the numerous specific images became too hard to understand, complex to maintain, and cumbersome to customise.
As part of the simplification, a recent innovation is to produce a `bioconductor_docker:devel` image to emulate the _Bioconductor_ nightly Linux build machine as closely as possible.
This image contains the build system environment variables and the _system dependencies_ needed to install and check almost all (1813 out of 1823) _Bioconductor_ software packages, and it saves users and package developers from managing these themselves.
Furthermore, the image is configured so that \code{.libPaths()} has \samp{/usr/local/lib/R/host-site-library} as the first location.
Users mounting a location on the host file system to this location can persistently manage installed packages across Docker containers or image updates.
Many R users pursue flexible work flows tailored to particular analysis needs rather than standardized work flows.
The new `bioconductor_docker` image is well suited for this preference, while `bioconductor_docker:devel` provides developers with a test environment close to _Bioconductor_'s build system.

@@

**\pkg{containerit}** ([https://github.com/o2r-project/containerit/](https://github.com/o2r-project/containerit/)) attempts to take this one step further and includes these tools to automatically create a `Dockerfile` that can execute a given workflow \citep{nust_containerit_2019}.
\pkg{containerit} accepts an R object of classes \code{"sessionInfo"} or \code{"session\_info"} as input and provides helper functions to derive these from workflows, e.g., an R script or R Markdown document, by analysing the session state at the end of the workflow.
It relies on the \pkg{sysreqs} ([https://github.com/r-hub/sysreqs/](https://github.com/r-hub/sysreqs/)) package and it's mapping of package system dependencies to platform specific installation package names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \CRANpkg{stevedore} to streamline the user interaction and improve the created `Dockerfile`s, e.g., by running a container for the desired base image to extract the already available R packages.

@@

**\CRANpkg{liftr}** aims to solve the problem of persistent reproducible reporting in statistical computing based on the R Markdown format \citep{xie2018} for dynamic documents \citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for collaborative documents and large-scale platforms for processing documents.
\CRANpkg{liftr} makes the document the main and sole workflow control file and the only file that needs to be shared between collaborators for consistent environments, e.g. demonstrated in the DockFlow project (https://dockflow.org).
It introduces new fields to the R Markdown document header, allowing users to manually declare the dependencies, including versions, for rendering the document.
The package then generates a `Dockerfile` from this metadata and provides a utitility function to render the document inside a Docker container, i.e., \code{render\_docker("foo.Rmd")}.
An RStudio addin even allows compilation of documents with the single push of a button.

@@

**\CRANpkg{googleComputeEngineR}** has the function \code{gce\_vm\_cluster()} to create clusters of 2 or more virtual machines, running multi-CPU architectures \citep{googleComputeEngineR_2019}.

@@

Furthermore, \CRANpkg{drake} workflows have been shown to use \CRANpkg{future} package's function \code{makeClusterPSOCK()} for sending parts of the workflow to a Docker image for execution\footnote{\href{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}} \citep[see package's function documentation;~][]{future_2020}.

@@

Another example is the package \CRANpkg{golem}, which makes an heavy use of \pkg{dockerfiler} when it comes to creating the `Dockerfile` for building and deploying production-grade Shiny applications \citep{cran_golem}.
\CRANpkg{googleComputeEngineR} enables quick deployments of key R services, such as RStudio and Shiny, onto cloud virtual machines (VMs) with Google Cloud Compute Engine \citep{googleComputeEngineR_2019}.
The package utilises `Dockerfile`s to move the labour of setting up those services from the user to a premade Docker image, which is configured and run in the cloud VM.
For example, by specifying the template `template="rstudio"` in functions \code{gce\_vm\_template()} and \code{gce\_vm()} an up-to-date RStudio Server image is launched for development work, whereas specifying `template="rstudio-gpu"` will launch an RStudio Server image with a GPU attached, etc.

### 2020-03-03 rockerverse.tex
move helper functions to README, add rockerverse.R
preparing for submission, see #26

also add missing used packages to DESCRIPTION
use \code{..} where appropriate, use DOI URLs
preparing for submission, #26

\textbf{\pkg{harbor}} (\url{https://github.com/wch/harbor}) is no longer
actively maintained, but it should be honorably mentioned as the first R
package for managing Docker images and containers. It uses the
\CRANpkg{sys} package \citep{cran_sys} to run system commands against
the Docker CLI, both locally and through an SSH connection, and it has
convenience functions, e.g., for listing and removing containers/images
and for accessing logs. The outputs of container executions are
converted to appropriate R types. The Docker CLI's basic functionality,
although it evolves quickly and with little concern for avoiding
breaking changes, has remained unchanged in core functions, meaning that
a core function such as
\texttt{harbor::docker\_run(image\ =\ "hello-world")} still works
depsite its stopped development.

\textbf{\CRANpkg{stevedore}} is currently the most powerful Docker
client in R \citep{cran_stevedore}. It interfaces with the Docker daemon
over the Docker HTTP
API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}}
via a Unix socket on Linux or MacOS, over a named pipe on Windows, or
over an HTTP/TCP connection. The package is the only one not using
system calls to the \texttt{docker} CLI tool for managing images and
containers. The package thereby enables connections to remote Docker
instances without direct configuration of the local Docker daemon.
Furthermore using the API gives access to information in a structured
way, is system independent, and is likely more reliable than parsing
command line output. \CRANpkg{stevedore}'s own interface is
automatically generated based on the OpenAPI specification of the Docker
daemon, but it is still similar to the Docker CLI. The interface is
similar to R6 objects, in that a \texttt{stevedore\_object} has a number
of functions attached to it that can be called, and multiple specific
versions of the Docker API can be supported thanks to the automatic
generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

@@

\textbf{\pkg{dockyard}}
(\url{https://github.com/thebioengineer/dockyard}) has the goal of
lowering the barrier to creating \texttt{Dockerfile}s, building Docker
images, and deploying Docker containers. The package follows the
increasingly used piping paradigm of the \texttt{tidyverse} style of
programming for chaining R functions representing the instructions in a
\texttt{Dockerfile}. An existing \texttt{Dockerfile} can be used as a
template. \pkg{dockyard} also includes wrappers for common steps, such
as installing an R package or copying files, and provides built-in
functions for building an image and running a container, which make
Docker more approachable within a single R-based user interface.

@@

The \emph{Bioconductor} \texttt{Dockerfile} selects the desired version
of R from Rocker, adds required system dependencies, and uses the
\CRANpkg{BiocManager} package for installing appropriate versions of
\emph{Bioconductor} packages \citep{cran_biocmanager}. A strength of
this approach is that the responsibility for complex software
configuration (including customized development) is shifted from the
user to the experienced \emph{Bioconductor} core team. However, a recent
audit of the \emph{Bioconductor} image stack \texttt{Dockerfile} led to
the deprecation of several community-maintained images, because the
numerous specific images became too hard to understand, complex to
maintain, and cumbersome to customise. As part of the simplification, a
recent innovation is to produce a \texttt{bioconductor\_docker:devel}
image to emulate the \emph{Bioconductor} nightly Linux build machine as
closely as possible. This image contains the build system environment
variables and the \emph{system dependencies} needed to install and check
almost all (1813 out of 1823) \emph{Bioconductor} software packages, and
it saves users and package developers from managing these themselves.
Furthermore, the image is configured so that \texttt{.libPaths()} has
\texttt{/usr/local/lib/R/host-site-library} as the first location. Users
mounting a location on the host file system to this location can
persistently manage installed packages across Docker containers or image
updates. Many R users pursue flexible work flows tailored to particular
analysis needs rather than standardized work flows. The new
\texttt{bioconductor\_docker} image is well suited for this preference,
while \texttt{bioconductor\_docker:devel} provides developers with a
test environment close to \emph{Bioconductor}'s build system.

@@

\textbf{\pkg{containerit}}
(\url{https://github.com/o2r-project/containerit/}) attempts to take
this one step further and includes these tools to automatically create a
\texttt{Dockerfile} that can execute a given workflow
\citep{nust_containerit_2019}. \pkg{containerit} accepts R
\texttt{sessionInfo} objects as input and provides helper functions to
derive these from workflows, e.g., an R script or R Markdown document,
by analysing the session state at the end of the workflow. It relies on
the \pkg{sysreqs} (\url{https://github.com/r-hub/sysreqs/}) package and
it's mapping of package system dependencies to platform specific
installation package
names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \CRANpkg{stevedore} to streamline the user
interaction and improve the created \texttt{Dockerfile}s, e.g., by
running a container for the desired base image to extract the already
available R packages.

@@

\textbf{\CRANpkg{liftr}} aims to solve the problem of persistent
reproducible reporting in statistical computing based on the R Markdown
format \citep{xie2018} for dynamic documents
\citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for
collaborative documents and large-scale platforms for processing
documents. \CRANpkg{liftr} makes the document the main and sole workflow
control file and the only file that needs to be shared between
collaborators for consistent environments, e.g.~demonstrated in the
DockFlow project (\url{https://dockflow.org}). It introduces new fields
to the R Markdown document header, allowing users to manually declare
the dependencies, including versions, for rendering the document. The
package then generates a \texttt{Dockerfile} from this metadata and
provides a utitility function to render the document inside a Docker
container, i.e., \texttt{render\_docker("foo.Rmd")}. An RStudio addin
even allows compilation of documents with the single push of a button.

@@

function \texttt{gce\_vm\_cluster()} to create clusters of 2 or more

@@

\CRANpkg{future} package's function \texttt{makeClusterPSOCK()} for

@@

Another example is the package \CRANpkg{golem}, which makes an heavy use
of \texttt{dockerfiler} when it comes to creating the
\texttt{Dockerfile} for building and deploying production-grade Shiny
applications \citep{cran_golem}. \CRANpkg{googleComputeEngineR} enables
quick deployments of key R services, such as RStudio and Shiny, onto
cloud virtual machines (VMs) with Google Cloud Compute Engine
\citep{googleComputeEngineR_2019}. The package utilises
\texttt{Dockerfile}s to move the labour of setting up those services
from the user to a premade Docker image, which is configured and run in
the cloud VM. For example, by specifying the template
\texttt{template="rstudio"} in functions
\texttt{gce\_vm\_template()/gce\_vm()} an up-to-date RStudio Server
image is launched for development work, whereas specifying
\texttt{template="rstudio-gpu"} will launch an RStudio Server image with
a GPU attached, etc.

>>>>

\textbf{\pkg{harbor}} (\url{https://github.com/wch/harbor}) is no longer
actively maintained, but it should be honorably mentioned as the first R
package for managing Docker images and containers. It uses the
\CRANpkg{sys} package \citep{cran_sys} to run system commands against
the Docker CLI, both locally and through an SSH connection, and it has
convenience functions, e.g., for listing and removing containers/images
and for accessing logs. The outputs of container executions are
converted to appropriate R types. The Docker CLI's basic functionality,
although it evolves quickly and with little concern for avoiding
breaking changes, has remained unchanged in core functions, meaning that
a core function such as
\code{harbor::docker\_run(image = "hello-world")} still works depsite
its stopped development.

\textbf{\CRANpkg{stevedore}} is currently the most powerful Docker
client in R \citep{cran_stevedore}. It interfaces with the Docker daemon
over the Docker HTTP
API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}}
via a Unix socket on Linux or MacOS, over a named pipe on Windows, or
over an HTTP/TCP connection. The package is the only one not using
system calls to the \texttt{docker} CLI tool for managing images and
containers. The package thereby enables connections to remote Docker
instances without direct configuration of the local Docker daemon.
Furthermore using the API gives access to information in a structured
way, is system independent, and is likely more reliable than parsing
command line output. \CRANpkg{stevedore}'s own interface is
automatically generated based on the OpenAPI specification of the Docker
daemon, but it is still similar to the Docker CLI. The interface is
similar to R6 objects, in that an object of class
\code{"stevedore\_object"} has a number of functions attached to it that
can be called, and multiple specific versions of the Docker API can be
supported thanks to the automatic
generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

@@

\textbf{\pkg{dockyard}}
(\url{https://github.com/thebioengineer/dockyard}) has the goal of
lowering the barrier to creating \texttt{Dockerfile}s, building Docker
images, and deploying Docker containers. The package follows the
increasingly used piping paradigm of the Tidyverse-style
\citep{wickham_welcome_2019} of programming for chaining R functions
representing the instructions in a \texttt{Dockerfile}. An existing
\texttt{Dockerfile} can be used as a template. \pkg{dockyard} also
includes wrappers for common steps, such as installing an R package or
copying files, and provides built-in functions for building an image and
running a container, which make Docker more approachable within a single
R-based user interface.

@@

The \emph{Bioconductor} \texttt{Dockerfile} selects the desired version
of R from Rocker, adds required system dependencies, and uses the
\CRANpkg{BiocManager} package for installing appropriate versions of
\emph{Bioconductor} packages \citep{cran_biocmanager}. A strength of
this approach is that the responsibility for complex software
configuration (including customized development) is shifted from the
user to the experienced \emph{Bioconductor} core team. However, a recent
audit of the \emph{Bioconductor} image stack \texttt{Dockerfile} led to
the deprecation of several community-maintained images, because the
numerous specific images became too hard to understand, complex to
maintain, and cumbersome to customise. As part of the simplification, a
recent innovation is to produce a \texttt{bioconductor\_docker:devel}
image to emulate the \emph{Bioconductor} nightly Linux build machine as
closely as possible. This image contains the build system environment
variables and the \emph{system dependencies} needed to install and check
almost all (1813 out of 1823) \emph{Bioconductor} software packages, and
it saves users and package developers from managing these themselves.
Furthermore, the image is configured so that \code{.libPaths()} has
\samp{/usr/local/lib/R/host-site-library} as the first location. Users
mounting a location on the host file system to this location can
persistently manage installed packages across Docker containers or image
updates. Many R users pursue flexible work flows tailored to particular
analysis needs rather than standardized work flows. The new
\texttt{bioconductor\_docker} image is well suited for this preference,
while \texttt{bioconductor\_docker:devel} provides developers with a
test environment close to \emph{Bioconductor}'s build system.

@@

\textbf{\pkg{containerit}}
(\url{https://github.com/o2r-project/containerit/}) attempts to take
this one step further and includes these tools to automatically create a
\texttt{Dockerfile} that can execute a given workflow
\citep{nust_containerit_2019}. \pkg{containerit} accepts an R object of
classes \code{"sessionInfo"} or \code{"session\_info"} as input and
provides helper functions to derive these from workflows, e.g., an R
script or R Markdown document, by analysing the session state at the end
of the workflow. It relies on the \pkg{sysreqs}
(\url{https://github.com/r-hub/sysreqs/}) package and it's mapping of
package system dependencies to platform specific installation package
names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \CRANpkg{stevedore} to streamline the user
interaction and improve the created \texttt{Dockerfile}s, e.g., by
running a container for the desired base image to extract the already
available R packages.

@@

\textbf{\CRANpkg{liftr}} aims to solve the problem of persistent
reproducible reporting in statistical computing based on the R Markdown
format \citep{xie2018} for dynamic documents
\citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for
collaborative documents and large-scale platforms for processing
documents. \CRANpkg{liftr} makes the document the main and sole workflow
control file and the only file that needs to be shared between
collaborators for consistent environments, e.g.~demonstrated in the
DockFlow project (\url{https://dockflow.org}). It introduces new fields
to the R Markdown document header, allowing users to manually declare
the dependencies, including versions, for rendering the document. The
package then generates a \texttt{Dockerfile} from this metadata and
provides a utitility function to render the document inside a Docker
container, i.e., \code{render\_docker("foo.Rmd")}. An RStudio addin even
allows compilation of documents with the single push of a button.

@@

function \code{gce\_vm\_cluster()} to create clusters of 2 or more

@@

\CRANpkg{future} package's function \code{makeClusterPSOCK()} for

@@

Another example is the package \CRANpkg{golem}, which makes an heavy use
of \pkg{dockerfiler} when it comes to creating the \texttt{Dockerfile}
for building and deploying production-grade Shiny applications
\citep{cran_golem}. \CRANpkg{googleComputeEngineR} enables quick
deployments of key R services, such as RStudio and Shiny, onto cloud
virtual machines (VMs) with Google Cloud Compute Engine
\citep{googleComputeEngineR_2019}. The package utilises
\texttt{Dockerfile}s to move the labour of setting up those services
from the user to a premade Docker image, which is configured and run in
the cloud VM. For example, by specifying the template
\texttt{template="rstudio"} in functions \code{gce\_vm\_template()} and
\code{gce\_vm()} an up-to-date RStudio Server image is launched for
development work, whereas specifying \texttt{template="rstudio-gpu"}
will launch an RStudio Server image with a GPU attached, etc.
